Regression Cube Analysis of Population Study Data

### 2015-06-22 draft.tex
added Bernhards revision of the cover letter
renamed heat map in heatmaps

%% Abstract section.
\abstract{%
%\com{Problem}.
Epidemiological studies comprise heterogeneous data about a subject group to define disease-specific risk factors.
%%
These data contain information (\emph{features}) about a subject's lifestyle, medical condition as well as medical image data.
%%
Statistical regression analysis is used to evaluate these features and to identify feature combinations indicating a disease (the \emph{target feature}).
%%
Although there is a strong demand for overview visualizations of a whole data set w.r.t. a target feature, no suitable tool is available for epidemiological researchers.

@@

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
%%
They extract classification rules that serve as basis for our proof-of-concept tests.
%%
Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
%%
\paragraph{Prior Work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%%
%We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%%
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%
We defined an Interactive Visual Analysis workflow for image-centric population study data \cite{Klemm2014VIS}.
%%
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
%%
Information visualizations of population study features were augmented with extracted image data, yielding linked views with both image and non-image information.
%%
The mosaic plot \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
%%
Its great popularity among our domain experts was the inspiration for this work.
%%

@@

The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
%%
}

@@

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes}.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
\noindent Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

@@

\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D mosaic plot of the slice.

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{Regression Cube} for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
\design{
%%
Early analysis sessions yielded two approaches on analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis is well suited for deriving new insight into the data, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
}
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

 (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.

@@

	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].

@@

\paragraph{Usage of a Regression Prism for Information Reduction.}
%%
Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the mosaic plot matrix.
%%
This is due to the symmetry of basic regression operators.
%%
$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
%%
Therefore, we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
%%
This opens up space for displaying additional information.
%%
Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
%%
The diagonal therefore acts as reference on how strong the correlation for the given row (or column) feature is.

@@

\paragraph{3D Prism as Data Mini-Map.}
\design{
%%
In early prototype versions, the 3D prism acted as starting point for the data analysis without the implementation of a separate 2D view.
%%
Slices were shown using cutaway planes.
%%
This approach was not popular with our domain experts, because the complexity of the visualization overwhelmed them. %did not like the combination of 3D interaction and slicing with representatives.
%%
}
%%
The 3D cube representation \design{was redesigned to} act as an overview over the whole data set.
%%
Its purpose is not to derive detailed information about data points.
%%
It serves as a function similar to a mini-map, guiding the attention to points of interest in the data.
%%
It also gives context information about adjacent data values when using the 2D mosaic plot.
%%
\design{The distinction between overview and details and demand using two different representations was well received with our domain experts.}
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes the absolute regression metric values (Fig.~\ref{fig:System}b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its regression metric values and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
\magenta{Variables are ordered the same way in the 2D mosaic plot and 3D cube to preserve the mental model and make them visually analogous.}

@@

	\textbf{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
	%%
%\end{itemize}
%%
The currently selected slice is denoted with a semi-transparent gray plane.
%%
\design{
Early designs rendered the whole cube, which made it hard to assess the position of the plane.
%%
%Since the regression metrics are mirrored along the diagonal, we decided to use this space to support better identification of the current slice.
Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
}
%%
The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
%%
The \magenta{regression metric values} are projected on this plane to provide an overlapping-free view.
%%
This allows for a easier to identify the current slice.
%%
\paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
%%
The 2D mosaic visualization (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice.
%%
It creates a optical equivalence with the 3D cube.
%%
To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation-based feature selection.
%%
The free space above the matrix diagonal is used to display the 3D cube.
%%

@@

%\paragraph{Front-End.}
%%
The front-end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
%%
\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
%%
It forces developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
%%
The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements with proper stylings.
%%
The 2D mosaic plot is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
%%
It provides fast and easy methods for binding data to graphical elements.
%%
The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
%%
% We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
% %%
% Open source volume rendering methods are available but do not satisfy our requirements.
% %%
% Creating a cube primitive for each data point resulted in non-interactive frame-rates for data sets larger than 30 features (creating 30$^3$ cube primitives).
% %%
% Therefore, we decided to use a shader-based solution by rendering the cube as a sprite-based particle system, allowing to customize color and opacity of every data point.
% %%
% It also is the fastest solution that we tested.

@@

%\paragraph{Access and Source.}
%%
A running instance of the \emph{Regression Cube} prototype can be found under \href{http://regressioncube.herokuapp.com/}{\texttt{regressioncube.herokuapp.com}}.
%%
The source for the prototype is freely available at \texttt{Github}.\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Node.js Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}
%%
Instructions and code on how to create a setup running the \emph{Regression Cube} statistical back-end through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
%%
The front-end can be deployed using \href{https://www.heroku.com/}{\texttt{www.heroku.com}} by cloning the repository into a Heroku app.
%%
\section{Application} \label{application}
%%
\begin{figure*}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/application}
 \caption{
 The analysis of the numerical and dichotomized target feature depicting liver fat values yields similar results (left).
 %%
 In (a), hotspots for somatometrical features with high correlations were found.
 %%
 High correlations were also found for features depicting \emph{hepatic steatosis} (b).
 %%
 A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D mosaic plot.
 %%
 The hypothesis-free analysis of the breast fat data set (right) w.r.t. the \emph{parenchyma tissue percentage} of the breast displays correlations between \emph{age}, \emph{body fat}, \emph{hip circumference} as well as \emph{menstrual period}.
 }
  \label{fig:Application}
\end{figure*}
%%
In this section, we describe how we applied the \emph{Regression Cube} to two epidemiological data sets.
%%
The hepatic steatosis data set was analyzed using data mining algorithms, yielding risk groups, which we now analyze further.
%%
Also, we try to reproduce the prior results from this analysis as proof-of-concept of our method.
%%
The breast fat data set is the basis for an explorative analysis w.r.t. the influencing parameters of the parenchyma tissue of the female breast.
%%

@@

%%
\textbf{Attention Steering is Crucial.}
%%
Important events have to be highlighted in overview visualizations to direct the user's attention to interesting parts of the data.
%%
Poor guidance potentially leads to overlooked relationships.
%%
We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose, e.g., for highlighting differences rather than displaying absolute values (Fig.~\ref{fig:Application}).
%%
\section{Summary and Outlook}
%%
We presented a technique for knowledge discovery in population study data sets with user-defined target features.
%%
Dimension reduction using the target restricts the analysis to the most important features.
%%
\emph{Hypothesis-free} analysis employs default regression models.
%%
Modeling expert knowledge using regression formulas allows for a \emph{hypothesis-based} investigation.
%%
A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
%%
These can then be analyzed further using the 2D mosaic plot for each cube slice.
%%
Details on demand for each model allow for a detailed assessment of regression models.
%%
We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
%%
The method was well received by our clinical partners, triggering detailed investigations of the findings.

>>>>

%% Abstract section.
\abstract{%
%\com{Problem}.
Epidemiological studies comprise heterogeneous data about a subject group to define disease-specific risk factors.
%%
These data contain information (\emph{features}) about a subject's lifestyle, medical status as well as medical image data.
%%
Statistical regression analysis is used to evaluate these features and to identify feature combinations indicating a disease (the \emph{target feature}).
%%
Although there is a strong demand for overview visualizations of a whole data set w.r.t. a target feature, no suitable tool is available for epidemiological researchers.

@@

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
%%
They extract classification rules that serve as basis for our proof-of-concept tests.
%%
Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
%%
\paragraph{Prior Work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%%
%We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%%
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%
We defined an Interactive Visual Analysis workflow for image-centric population study data \cite{Klemm2014VIS}.
%%
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
%%
Information visualizations of population study features were augmented with extracted image data, yielding linked views with both image and non-image information.
%%
The 2D heatmap \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
%%
Its great popularity among our domain experts was the inspiration for this work.
%%

@@

The idea emerged from analysis sessions of a previous project, which contained a 2D heatmap showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a 2D heatmap for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heatmap visualizations.
 %%
 These represent the slices in our cube.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heatmap.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heatmap by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
%%
}

@@

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes}.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D heatmap representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
\noindent Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

@@

\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D heatmap of the slice.

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D heatmap or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{Regression Cube} for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
\design{
%%
Early analysis sessions yielded two approaches on analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis is well suited for deriving new insight into the data, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
}
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

 (c) 2D heatmap visualization of the selected slice for feature \emph{Pain/Discomfort}.

@@

	\item The \emph{cube visualization} consisting of the 2D heatmap as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].

@@

\paragraph{Usage of a Regression Prism for Information Reduction.}
%%
Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the 2D heatmap matrix.
%%
This is due to the symmetry of basic regression operators.
%%
$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
%%
Therefore, we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
%%
This opens up space for displaying additional information.
%%
Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
%%
The diagonal therefore acts as reference on how strong the correlation for the given row (or column) feature is.

@@

\paragraph{3D Prism as Data Mini-Map.}
\design{
%%
In early prototype versions, the 3D prism acted as starting point for the data analysis without the implementation of a separate 2D view.
%%
Slices were shown using cutaway planes.
%%
This approach was not popular with our domain experts, because the complexity of the visualization overwhelmed them. %did not like the combination of 3D interaction and slicing with representatives.
%%
}
%%
The 3D cube representation \design{was redesigned to} act as an overview over the whole data set.
%%
Its purpose is not to derive detailed information about data points.
%%
It serves as a function similar to a mini-map, guiding the attention to points of interest in the data.
%%
It also gives context information about adjacent data values when using the 2D heatmap.
%%
\design{The distinction between overview and details and demand using two different representations was well received with our domain experts.}
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes the absolute regression metric values (Fig.~\ref{fig:System}b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its regression metric values and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
\magenta{Variables are ordered the same way in the 2D heatmap and 3D cube to preserve the mental model and make them visually analogous.}

@@

	\textbf{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
	%%
%\end{itemize}
%%
The currently selected slice is denoted with a semi-transparent gray plane.
%%
\design{
Early designs rendered the whole cube, which made it hard to assess the position of the plane.
%%
%Since the regression metrics are mirrored along the diagonal, we decided to use this space to support better identification of the current slice.
Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
}
%%
The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
%%
The \magenta{regression metric values} are projected on this plane to provide an overlapping-free view.
%%
This allows for a easier to identify the current slice.
%%
\paragraph{2D heatmap Plot Slice Visualization [4.DD].}
%%
The 2D heatmap (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice.
%%
It creates a optical equivalence with the 3D cube.
%%
To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation-based feature selection.
%%
The free space above the matrix diagonal is used to display the 3D cube.
%%

@@

%\paragraph{Front-End.}
%%
The front-end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
%%
\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
%%
It forces developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
%%
The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements with proper stylings.
%%
The 2D heatmap is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
%%
It provides fast and easy methods for binding data to graphical elements.
%%
The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
%%
% We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
% %%
% Open source volume rendering methods are available but do not satisfy our requirements.
% %%
% Creating a cube primitive for each data point resulted in non-interactive frame-rates for data sets larger than 30 features (creating 30$^3$ cube primitives).
% %%
% Therefore, we decided to use a shader-based solution by rendering the cube as a sprite-based particle system, allowing to customize color and opacity of every data point.
% %%
% It also is the fastest solution that we tested.

@@

%\paragraph{Access and Source.}
%%
A running instance of the \emph{Regression Cube} prototype can be found under \href{http://regressioncube.herokuapp.com/}{\texttt{regressioncube.herokuapp.com}}.
%%
The source for the prototype is freely available at \texttt{Github}.\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Node.js Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}
%%
Instructions and code on how to create a setup running the \emph{Regression Cube} statistical back-end through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
%%
The front-end can be deployed using \href{https://www.heroku.com/}{\texttt{www.heroku.com}} by cloning the repository into a Heroku app.
%%
\section{Application} \label{application}
%%
\begin{figure*}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/application}
 \caption{
 The analysis of the numerical and dichotomized target feature depicting liver fat values yields similar results (left).
 %%
 In (a), hotspots for somatometrical features with high correlations were found.
 %%
 High correlations were also found for features depicting \emph{hepatic steatosis} (b).
 %%
 A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D heatmap.
 %%
 The hypothesis-free analysis of the breast fat data set (right) w.r.t. the \emph{parenchyma tissue percentage} of the breast displays correlations between \emph{age}, \emph{body fat}, \emph{hip circumference} as well as \emph{menstrual period}.
 }
  \label{fig:Application}
\end{figure*}
%%
In this section, we describe how we applied the \emph{Regression Cube} to two epidemiological data sets.
%%
The hepatic steatosis data set was analyzed using data mining algorithms, yielding risk groups, which we now analyze further.
%%
Also, we try to reproduce the prior results from this analysis as proof-of-concept of our method.
%%
The breast fat data set is the basis for an explorative analysis w.r.t. the influencing parameters of the parenchyma tissue of the female breast.
%%

@@

%%
\textbf{Attention Steering is Crucial.}
%%
Important events have to be highlighted in overview visualizations to direct the user's attention to interesting parts of the data.
%%
Poor guidance potentially leads to overlooked relationships.
%%
We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose, e.g., for highlighting differences rather than displaying absolute values (Fig.~\ref{fig:Application}).
%%
\section{Summary and Outlook}
%%
We presented a technique for knowledge discovery in population study data sets with user-defined target features.
%%
Dimension reduction using the target restricts the analysis to the most important features.
%%
\emph{Hypothesis-free} analysis employs default regression models.
%%
Modeling expert knowledge using regression formulas allows for a \emph{hypothesis-based} investigation.
%%
A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
%%
These can then be analyzed further using the 2D heatmap for each cube slice.
%%
Details on demand for each model allow for a detailed assessment of regression models.
%%
We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
%%
The method was well received by our clinical partners, triggering detailed investigations of the findings.
